---
title: "study_usage_Sankey"
output: html_document
date: "2023-06-02"
---

Document for making Sankey Plot.

On the left side of the plot, we need to have the primary sources
On the right, we need to put how they are used in systematic reviews.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#See which libary we can load
```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
# if (!requireNamespace("ggsankey", quietly = TRUE))  install.packages("ggsankey")
# library("ggsankey")
```


```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("networkD3", quietly = TRUE))  BiocManager::install("networkD3")
# Load package
library(networkD3)
# Load energy projection data
URL <- paste0(
        "https://cdn.rawgit.com/christophergandrud/networkD3/",
        "master/JSONdata/energy.json")
Energy <- jsonlite::fromJSON(URL)
# Plot
sankeyNetwork(Links = Energy$links, Nodes = Energy$nodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             units = "TWh", fontSize = 12, nodeWidth = 30)
```

```{r}
study_usage <- read.table("study_update_KH_62323.txt", sep="\t", header = T,
                          encoding = "latin1", check.names = FALSE)
study_usage <- subset(study_usage, select = -`Reference Short`)
colnames(study_usage)[colnames(study_usage) == "Reference Short 2"] <- "Reference Short"
head(study_usage)
```
#Plan:
Need to programatically tease out node names c("RCT",Cross-sectional","Cohort", "MA", "T").
Need to find programatically linkages.
  Source and target
Need to make counts (values) for thickness of linkages.

```{r}
rows_use <- nrow(study_usage)-1#because the last line is different data
left_node_names <- unique(study_usage$`Study Design`[1:rows_use])
right_node_names <- c(colnames(study_usage)[4:ncol(study_usage)])
all_node_names <- c(left_node_names, right_node_names)

links <- data.frame(source = character(0), target = character(0), count = integer(0))
for (rw in 1:rows_use){
  primary <- study_usage$`Study Design`[rw]#source
  if (primary %in% left_node_names){
    used <- FALSE #flag for determining secondary usage
    source_node_num <- match(primary, all_node_names)
    # print(primary)
    for (cl in 4:ncol(study_usage)){
      query_cell <- study_usage[rw,cl]
      if (query_cell != "0"){
        used <- TRUE
        target_node_num <- match(colnames(study_usage)[cl], all_node_names)#match returns index
        #look up if source node and target node already in links
        current_link_combos <- paste(links$source, links$target)#get current combos
        if (paste(source_node_num, target_node_num) %in% current_link_combos){
          node_count <- links[links$source==source_node_num & links$target==target_node_num, "count"]
          # print(paste("final if:", source_node_num, target_node_num, node_count))
          links[links$source==source_node_num & links$target==target_node_num, "count"] <- node_count + 1
        }#if (in link_combos) close
        else{#add the node links if they aren't already in the df
          new_row <- data.frame("source" = source_node_num,"target" = target_node_num,"count" = 1)
          links <- rbind(links, new_row)
        }#else close
      }#if (query_cell %in% right_node_names close
    }#column for-loop close
    # if (used == FALSE){
    #   #if the primary source wasn't used, we increment the "None" count
    #   none_index <- match("None", all_node_names)
    #   current_link_combos <- paste(links$source, links$target)#get current combos
    #   if (paste(source_node_num, none_index) %in% current_link_combos){
    #       node_count <- links[links$source==source_node_num & links$target==none_index, "count"]
    #       # print(paste("'Used' if:", source_node_num, target_node_num, node_count))
    #       links[links$source==source_node_num & links$target==none_index, "count"] <- node_count + 1
    #     }#if (in link_combos) close
    #     else{
    #       new_row <- data.frame("source" = source_node_num,"target" = none_index,"count" = 1)
    #       links <- rbind(links, new_row)
    #     }#else close
    # }#if (used == FALSE
  }#if (primary) close
}#row for loop close

#update node names to clarify in plot


#make node names into dataframe for Sankey input
nodes <- data.frame(nodes = all_node_names)
#Adjust links because JS is 0 indexed
links$source <- links$source - 1
links$target <- links$target - 1

print(links)
```

```{r}
p <- sankeyNetwork(Links = links, Nodes = nodes, Source = "source",
             Target = "target", Value = "count", NodeID = "nodes",
             units = "Studies", fontSize = 20, nodeWidth = 30, 
             margin = list(50,100,500,100), sinksRight = TRUE,
             height = 500, width = 500)

saveNetwork(p, file="output/Study_Usage_Sankey_small_refs_KH20Jun2023.html", selfcontained = TRUE)
p

```

To improve the figure, Dr. Higgins has provided an updated and redesigned input and updated the request for the figure.


```{r}
rows_use <- nrow(study_usage)-1#because the last line is different data
ref_column_start <- 4
left_node_names <- unique(study_usage$`Reference Short`[1:rows_use])
right_node_names <- c(colnames(study_usage)[ref_column_start:ncol(study_usage)])
all_node_names <- c(left_node_names, right_node_names)
left_node_groups <- study_usage$`Study Design`[1:rows_use]
right_node_groups <- study_usage[nrow(study_usage),ref_column_start:ncol(study_usage)]
all_groups <- c(unlist(left_node_groups), unlist(right_node_groups))
length(all_node_names)
```

#Updated code to include node groupings
```{r}
links <- data.frame(source = character(0), target = character(0), count = integer(0))
# node_groups[length(all_node_names)] <- "None"
for (rw in 1:rows_use){
  primary <- study_usage$`Reference Short`[rw]#source
  source_node_num <- match(primary, all_node_names)
  # print(paste(rw, primary, study_usage$`Study Design`[rw]))
  for (cl in ref_column_start:ncol(study_usage)){
    query_cell <- study_usage[rw,cl]
    if (query_cell != "0"){
      target_node_num <- match(colnames(study_usage)[cl], all_node_names)#match returns index of matching var
      #look up if source node and target node already in links
      current_link_combos <- paste(links$source, links$target)#get current combos
      # print(current_link_combos)
      if (paste(source_node_num, target_node_num) %in% current_link_combos){
        node_count <- links[links$source==source_node_num & links$target==target_node_num, "count"]
        print(paste("final if:", source_node_num, target_node_num, node_count))
        links[links$source==source_node_num & links$target==target_node_num, "count"] <- node_count + 1
      }#if (in link_combos) close
      else{
        new_row <- data.frame("source" = source_node_num,"target" = target_node_num,"count" = 1)
        links <- rbind(links, new_row)
      }#else close
    }#if (query_cell %in% right_node_names close
  }#column for-loop close
  # if (target_group == "None"){
  #   #if the primary source wasn't used, we increment the "None" count
  #   none_index <- match("None", all_node_names)
  #   current_link_combos <- paste(links$source, links$target)#get current combos
  #   if (paste(source_node_num, none_index) %in% current_link_combos){
  #       node_count <- links[links$source==source_node_num & links$target==none_index, "count"]
  #       # print(paste("'Used' if:", source_node_num, target_node_num, node_count))
  #       links[links$source==source_node_num & links$target==none_index, "count"] <- node_count + 1
  #     }#if (in link_combos) close
  #     else{
  #       new_row <- data.frame("source" = source_node_num,"target" = none_index,"count" = 1)
  #       links <- rbind(links, new_row)
  #     }#else close
  # }#if (used == FALSE
}#row for loop close

#make node names into dataframe for Sankey input
nodes <- data.frame(nodes = all_node_names, groups=all_groups)
#Adjust links because JS is 0 indexed
links$source <- links$source - 1
links$target <- links$target - 1

# print(links)
```


#New Sankey network:
```{r}

# Add a 'group' column to the nodes data frame:# Add a 'group' column to the nodes data frame:
nodes$groups <- as.factor(nodes$groups)
nodes$groups <- gsub(" ", "_", nodes$groups)

links$groups <- lapply(links$source, function(x) {
  return(nodes$groups[x+1])#the "+1" is because the javascript is 0-indexed
  })
 
# Give a color for each group:
my_color <- 'd3.scaleOrdinal() .domain(["RCT", "Cross-sectional", "Cohort", "RCT_only", "Cohort_only", "RCT_and_Cohort"]) .range(["green", "grey", "red", "yellow", "orange", "blue"])'

p <- sankeyNetwork(Links = links, Nodes = nodes, Source = "source",
             Target = "target", Value = "count", NodeID = "nodes",
             fontSize = 10, nodeWidth = 40, NodeGroup = "groups",
             sinksRight = TRUE, nodePadding = 3, colourScale=my_color,
             LinkGroup="groups")
             #margin = list(50,100,500,100))
             #, height = 500, width = 500)

saveNetwork(p, file="output/no_none_study_Usage_Sankey.html", selfcontained = TRUE)

# render with js
sankey_rendered <- htmlwidgets::onRender(p,
  'function(el, x) {
    d3.selectAll(".node text")
        .style("fill", "blue");
  }'
)

sankey_rendered
saveNetwork(sankey_rendered, file="output/no_none_blue_study_Usage_Sankey.html",selfcontained = TRUE)

```
#Simple network plot





```{r}
p <- networkD3::simpleNetwork(links)

p
```

```{r}
p <- networkD3::forceNetwork(Links = links, Nodes = nodes, Source = "source",
             Target = "target", Value = "count", NodeID = "nodes", Group = "groups",
             fontSize = 9)

p
```





https://stackoverflow.com/questions/35056733/how-to-capture-html-output-as-png-in-r

http://palomnyk.github.io/Sankey/output/Study_Usage_Sankey.html


