---
title: "study_usage_Sankey"
output: html_document
date: "2023-06-02"
---

Document for making Sankey Plot.

On the left side of the plot, we need to have the primary sources
On the right, we need to put how they are used in systematic reviews.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
study_usage <- read.table("study_usage.txt", sep="\t", header = T,
                          encoding = "latin1")
head(study_usage)
```

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
# if (!requireNamespace("ggsankey", quietly = TRUE))  BiocManager::install("ggsankey")
# library("ggsankey")
# install.packages("remotes")
# remotes::install_github("NumbersInternational/flipPlots")

# install.packages("flipplots")
```






```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("networkD3", quietly = TRUE))  BiocManager::install("networkD3")
# Load package
library(networkD3)
# Load energy projection data
URL <- paste0(
        "https://cdn.rawgit.com/christophergandrud/networkD3/",
        "master/JSONdata/energy.json")
Energy <- jsonlite::fromJSON(URL)
# Plot
sankeyNetwork(Links = Energy$links, Nodes = Energy$nodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             units = "TWh", fontSize = 12, nodeWidth = 30)
```

#Plan:
Need to programatically tease out node names c("RCT",Cross-sectional","Cohort", "MA", "T").
Need to find programatically linkages.
  Source and target
Need to make counts (values) for thickness of linkages.

#Psudo-code to find linkages and counts
links <- dataframe for links with headers as: source, target, count
For row in study_usage, if study_usage$Study.Design is present:
  used <- FALSE #flag for determining secondary usage
  source_node_number <- index in all node names
  for column in secondary columns:
    if row in column has name in node_names:
      used <- TRUE
      target_node_number <- index in all node names
      if source_node_number and target_node_number combo in linkds df:
        links[source == source_node_number & target == target_node_number, counts] += 1
      else:
        links[source == source_node_number & target == target_node_number, counts] <- 1
    if (used == FALSE):
        links[source == source_node_number & target == "None", counts] += 1
      else:
        links[source == source_node_number & target == "None", counts] <- 1

```{r}
left_node_names <- unique(study_usage$Study.Design)
right_node_names <- c("T","MA","None")
all_node_names <- c(left_node_names, right_node_names)

links <- data.frame(source = character(0), target = character(0), count = integer(0))
for (rw in 1:nrow(study_usage)){
  primary <- study_usage$Study.Design[rw]#source
  if (primary %in% left_node_names){
    used <- FALSE #flag for determining secondary usage
    source_node_num <- match(primary, all_node_names)
    # print(primary)
    for (cl in 10:19){
      query_cell <- study_usage[rw,cl]
      if (query_cell %in% right_node_names){
        used <- TRUE
        target_node_num <- match(query_cell, all_node_names)
        #look up if source node and target node already in links
        current_link_combos <- paste(links$source, links$target)#get current combos
        if (paste(source_node_num, target_node_num) %in% current_link_combos){
          node_count <- links[links$source==source_node_num & links$target==target_node_num, "count"]
          # print(paste("final if:", source_node_num, target_node_num, node_count))
          links[links$source==source_node_num & links$target==target_node_num, "count"] <- node_count + 1
        }#if (in link_combos) close
        else{
          new_row <- data.frame("source" = source_node_num,"target" = target_node_num,"count" = 1)
          links <- rbind(links, new_row)
        }#else close
      }#if (query_cell %in% right_node_names close
    }#column for-loop close
    if (used == FALSE){
      #if the primary source wasn't used, we increment the "None" count
      none_index <- match("None", all_node_names)
      current_link_combos <- paste(links$source, links$target)#get current combos
      if (paste(source_node_num, none_index) %in% current_link_combos){
          node_count <- links[links$source==source_node_num & links$target==none_index, "count"]
          # print(paste("'Used' if:", source_node_num, target_node_num, node_count))
          links[links$source==source_node_num & links$target==none_index, "count"] <- node_count + 1
        }#if (in link_combos) close
        else{
          new_row <- data.frame("source" = source_node_num,"target" = none_index,"count" = 1)
          links <- rbind(links, new_row)
        }#else close
    }#if (used == FALSE
  }#if (primary) close
}#row for loop close

#update node names to clarify in plot
all_node_names[all_node_names == "MA"] <- "Meta-analysis"
all_node_names[all_node_names == "T"] <- "Text/tabular summary"
#make node names into dataframe for Sankey input
nodes <- data.frame(nodes = all_node_names)
#Adjust links because JS is 0 indexed
links$source <- links$source - 1
links$target <- links$target - 1

print(links)
```

```{r}

p <- sankeyNetwork(Links = links, Nodes = nodes, Source = "source",
             Target = "target", Value = "count", NodeID = "nodes",
             units = "Studies", fontSize = 20, nodeWidth = 30, 
             margin = list(50,100,500,100), sinksRight = TRUE,
             height = 500, width = 500)

saveNetwork(p, file="output/Study_Usage_Sankey.html", selfcontained = TRUE)
p

```


I want to customize this more than NetworkD3 allows, so now add column names:
Getting the code from here:
https://stackoverflow.com/questions/66813278/how-to-add-columnn-titles-in-a-sankey-chart-networkd3


```{r}

htmlwidgets::onRender(p, '
  function(el) { 
    var cols_x = this.sankey.nodes().map(d => d.x).filter((v, i, a) => a.indexOf(v) === i).sort(function(a, b){return a - b});
    var labels = ["Primary Studies", "Secondary Studies"];
    cols_x.forEach((d, i) => {
      d3.select(el).select("svg")
        .append("text")
        .attr("x", d)
        .attr("y", 12)
        .text(labels[i]);
    })
  }
')

htmlwidgets::saveWidget(
  p,
  "output/updated_Sankey.html",
  selfcontained = TRUE,
)

```

I also want to add a label:
https://www.geeksforgeeks.org/how-to-add-axis-labels-using-networkd3-in-r/

```{r}


p <- htmlwidgets::prependContent(p,
                                 htmltools::tags$h1(
                                 "Sweetness Primary Study Usage",
                                 style="text-align:center;color:#000000; "),
                                      htmltools::tags$h2(
                                        "Sankey plot",
                                      style="color:#000000;text-align:center;"))
 
# view graph
htmlwidgets::saveWidget(
  p,
  "output/updated_Sankey.html",
  selfcontained = TRUE
)
p

```



```{r}

# create left-right label
leftTx = htmltools::tags$div(
         style="max-width: 30vw; height: 100%;
                 display: flex; align-items: center;
                 justify-content: center;",
         tags$p(""))
 
rightTx = htmltools::tags$p("",
                   style="max-width:30vw")
 
# final output with label. combineWidget
# can be use by manipulateWidget.
p <- combineWidgets(p,
                    title = tags$h1("With left-right label",
                    style="text-align:center;color:#007500;"),
                    leftCol = leftTx,
                    rightCol = rightTx,
                    nrow = 1)
 
# view graph
p

```




# create left-right label
leftTx = tags$div(
         style="max-width: 30vw; height: 100%;
                 display: flex; align-items: center;
                 justify-content: center;",
         tags$p("This is left side label"))
 
rightTx = tags$p("This is right side label",
                   style="max-width:30vw")
 
# final output with label. combineWidget
# can be use by manipulateWidget.
graph3 <- combineWidgets(graph2,
                      title = tags$h1("With left-right label",
                      style="text-align:center;color:#007500;"),
                      leftCol = leftTx,
                      rightCol = rightTx,
                      nrow = 1)
 
# view graph
graph3

https://stackoverflow.com/questions/35056733/how-to-capture-html-output-as-png-in-r




